# Data and methods

## Environmental Patenting in OECD countries

To measure the drivers of environmental innovation I rely on the NIC
approach described above, and first introduced by
@furmanUnderstandingDriversNational2000. However, I use a modified
version, taking insights mainly from
@andrijauskieneRedevelopingNationalInnovative2021. Those authors
redevelop the original NIC framework in a broader fashion and apply it
to European countries. However, their framework is only appropriate for
the national and not the sectoral or technology-specific fields. In
particular on environmental technologies, I follow an approach similar
to that of @moroNationalInnovativeCapacity2019. They also use a modified
version of the NIC approach to study the drivers of innovative outputs
in the water sector in the European Union and China. The main difference
with this thesis is that the range of technologies included and the
selected sample.

Empirically, this thesis uses a panel data set of 36 member countries
from OECD from 2000 to 2020. The study focuses on OECD countries due to
the availability of data, and also their comparability in terms of
similar levels of economic development and their adherence to OECD
standards of environmental protection. On the other hand, OECD includes
a diverse set of countries in terms of population, size, technological
specialization, and geographical representation, giving a sample with
enough variation to yield substantive results.

It is worth noting that not all current OECD countries are included in
the sample. Costa Rica and Colombia, now full members of the OECD,
joined the after the dates of interest of this work, in 2020 and 2021
respectively, and are thus excluded from the analysis.

The main dependent variable is the number of green patents per million
inhabitants produced by country. Patents have long been used as a
measure of the output of innovative activity. One of the first to point
out the advanges of working with patents was
@grilichesPatentStatisticsEconomic1990, who mentions that "they are
available; they are by definition related to inventiveness, and they are
based on what appears to be an objective and only slowly changing
standard". Another advantage pointed out by
@hascicMeasuringEnvironmentalInnovation2015 is their comparability,
which make them specially suited for cross-country comparisons like the
one carried out in this thesis.

That does not mean that patents come without drawbacks. For instance,
patents do not cover the full spectrum of the innovative output; they
mostly do not capture process innovation, and social innovation, for
instance. Another drawback is that patents do not account for the
quality of inventions or their diffusion. Indeed, it is known that the
economic value of patents is very skewed, with a few of them worth a
lot; and most of them with very low economic value
[@hascicMeasuringEnvironmentalInnovation2015].

Nonetheless, for the moment there do not seem to be a better indicator
of output of innovative activity, and therefore I rely on the data
provided by the green patent database of the OECD to build the dependent
variable. The count includes four main categories of technologies, i.e.
environmental management, climate change mitigation, climate change
adaptation, and sustainable ocean economy.
Appendix XX shows the patent classes included in the dependent variable.

It is important to mention that working with patent data has its challenges. Achieving comparability across countries and representativeness  within a country requires familiarity with the way patents work and how patent statistics are collected [@hascicUsePatentStatistics2015]. Especially when dealing with narrow technological fields, the researcher must make methodological decisions regarding what exactly to include in its data. Patents are classified by classes that form technology domains, and family size.

The family size refers to a collection of patent applications covering the same or similar technical content. The applications in a family are related to each other through priority claims. For this work, I use a "patent family >= 2", meaning that the invention has been claimed in at least two different jurisdiccions. @hascicUsePatentStatistics2015 mentions that “In general, PF2 seems as a suitable compromise between comparability (quality threshold) and practicality (variance), except for very narrow technological fields or small inventor countries – in which case PF1 might be the only option [...] It turns out that for many applications related to environmental technologies, the PF2 restriction (claimed priorities) is particularly suitable in order to achieve the dual purpose of comparability and robustness. The count of claimed priorities (i.e. patent applications deposited at any office world-wide, that have been claimed as priority elsewhere in the world) is a suitable statistic for the purpose of international comparisons because only the priority applications protecting “high-value” invention are counted". Since I will be working a variety of countries in a not-very-narrow technological field, it is reasonable to use PF2. The patent is assigned to a country according to the place of residence of the inventor (as opposed to the nationality, for example). When a patent names inventors residing in different countries, each country is assigned a fractional count of the patent. For instance when a patent is the result of two inventors residing in different countries, each country will be given a count of 0.5, and so on for more countries. This is why the variable takes a continuous instead of discrete value.

It is common knowledge that the production of technology is general, or at least of patented
technology is very concentrated in a few countries and regions
[@oecdOECDPatentStatistics2009]. There are various reasons for that.
First, a large part of the innovative activity happen in big
laboratories located in a few countries (and even a few regions) of the
world. Second, there are differences in the patenting propensity of
firms, given the different regulations and the costs associated to
patenting in different jurisdictions.

In general, the OECD countries included in this thesis represent
with some of the world's highest innovative activity and technological
endowment. Indeed, 17 out the 20 countries that filed more applications in 2020
were members of the OECD (the other three were China, India and Russia).
However, patenting activity has shifted in the last decade. China has
taken a more important role, both in terms of origin of the patents
(that is, the country of residency of the fist inventor named in the
patent filing) and the number of filings in the Chinese patent office
(CPA), both of which accounted for roughly 50% of the world's patenting
activity in 2020 [@worldintellectualpropertyorganization.WorldIntellectualProperty2021].

However, even in that sample the distribution is quite
skewed, with some countries concentrating most of the production of
technologies and others with very little patenting activity. Chart \@ref(fig:gp-by-country) shows the patenting activity in
environmental technologies of the OECD countries from 2000 to 2019. This figure shows that most of the patenting in environmental
technologies happen in six countries, the United States, Japan, Korea,
Germany, France and the United Kingdom. Out of those, the United States, Japan, ans in some years Korea, generate on average more patents than the 30 OECD countries with less patenting taken together (Others). We can also see that, in general, there was a sharp increase in the production of green patents since the beginning of the century, and especially since 2005 up until around 2012, when the tendency changes and most countries start decreasing their annual production of green patented technologies. 

```{r, include FALSE, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
library(plotly)
library(plm)
library(pglm)
library(panelr)
library(skimr)
library(ggthemes)
library(gridExtra)
library(stargazer)
library(modelsummary)
library(kableExtra)
library(gtsummary)
library(janitor)
library(corrplot)
library(knitr)
library(tinytex)
library(huxtable)
library(vtable)
library(cowplot)
library(treemap)
library(hrbrthemes)

###################################################################

df2 <- read.csv("G:/My Drive/EPOG/Thesis/Green_National_Innovative_Capacity/df2.csv")

df4 <- df2 %>% select(
  Country, Year,
  green_patents, gp_mill, 
  gerd_ppp, rd_tt_pt, ter_edu_gdp,  gerd_per_hes,  gerd_fin_bs,
  stock_gp, epsi, green_taxes, env_rd_totalrd, 
  reg_quality, cont_corr)

descriptive_stats <- as.data.frame(df4) %>% select(
  green_patents, gp_mill, 
  gerd_ppp, rd_tt_pt, ter_edu_gdp,  gerd_per_hes,  gerd_fin_bs,
  stock_gp, epsi, green_taxes, env_rd_totalrd, 
  reg_quality, cont_corr)
```

```{r Fig_1_code, echo=FALSE, include=FALSE}
# 1. Green patents by country

sum_others <- df2 %>% filter(!Country %in% c("United States", "Korea", "Japan", "Germany", "France", "United Kingdom")) %>% group_by(Year) %>% filter(!Year %in% c(2020)) %>%  summarise(sum = sum(green_patents, na.rm = TRUE)) %>% mutate(Country = "Others")

gp_by_country <- df2 %>%
  filter(Country %in% c("United States", "Korea", "Japan", "Germany", "France", "United Kingdom")) %>% 
  ggplot(mapping = aes(x = Year, y = green_patents, color = Country)) +
  geom_line(linewidth = 1.2) +
  geom_line(data = sum_others, aes (x = Year, y = sum, color = Country), linewidth = 1.2) +
  labs(title = "Green patents in OECD countries 2000-2019", y = "Number of green patents", x = "Year") +
  theme_clean()

## 2. Green patents per capita

gp_per_capita <- df2 %>%
  filter(Country %in% c("United States", "Korea", "Japan", "Germany", "France", "United Kingdom")) %>% 
  ggplot(mapping = aes(x = Year, y = gp_mill, color = Country)) +
  geom_line(linewidth = 1.2) +
  labs(title = "Green patents per million population
       in OECD countries 2000-2019", y = "Number of green patents", x = "Year") +
  theme_clean()

## 3. Green patents by unit of GDP

gp_per_gdp <- df2 %>%
  filter(Country %in% c("United States", "Korea", "Japan", "Germany", "France", "United Kingdom")) %>% 
  mutate(patents_gdp = green_patents / GDP) %>% 
  ggplot(mapping = aes(x = Year, y = patents_gdp, color = Country)) +
  geom_line(linewidth = 1.2) +
  labs(title = "Green patents per unit of GDP", y = "Number of green patents", x = "Year") +
  theme_clean()

```

```{r gp-by-country, warning=FALSE, echo=FALSE}
#| label: gp-by-country
#| fig-cap: Green technology patenting by country
gp_by_country
```


Looking a bit further into the six countries that produce more green patents, Fig. \@ref(fig:green-patents-productivity) shows their productivity, measured as number of green patents per million population and per unit of GDP. In both cases, we can see that Korea, Japan and Germany are the countries with the highest productivity. Korea in particular, shifted from the other countries around 2007 and became more productive by an important margin and keeping this advantage until now, even if some decline can be observed in the last years.


```{r green-patents-productivity, echo=FALSE, warning=FALSE}
#| label: green-patents-productivity
#| fig-cap: Green patent productivity

grid.arrange(gp_per_capita, gp_per_gdp)
```

These environmental patents are not evenly distributed across technology domains.
Indeed, climate change mitigation technologies account for most of the
innovation during the whole period of study (59%). As shown in Chart \@ref(fig:green-patents-domain), since 2000, this type of
technologies increased sharply and only had a declining trend from 2012,
but still keeping as the biggest group by a large margin. Environmental
management technologies also increased particularly from 2003 until
2010; after that the tendency becomes less consistent but generally
downward. In contrast, climate change adaptation and sustainable ocean
technologies have been quite stable during the whole period, accounting
for around 8.5% and 2.5% respectively during the two-decade period of analysis.

```{r green_patents_domain_code, echo=FALSE, include=FALSE}
Green_patents_OECD2 <- read_csv("G:/My Drive/EPOG/Thesis/Green_National_Innovative_Capacity/env_pat_1995_2019_by_tech_domains.csv")
all_tech <- Green_patents_OECD2 %>% filter(`Technology domain` == "All technologies (total patents)")
gp2 <- Green_patents_OECD2 %>% filter (!`Technology domain` == "All technologies (total patents)" & !`Technology domain` == "Environment-related technologies")

gp2$Year <- as.Date((as.character(gp2$Year)), format = "%Y")

gp2 <- gp2 %>% rename(tech_dom = `Technology domain`)

gp_by_tech_dom <- gp2 %>% group_by(Year, tech_dom) %>% summarise(sum = sum(Value)) %>% 
  ggplot(mapping = aes(x = Year, y = sum)) +
  geom_line(aes(color = tech_dom), linewidth = 1.2) +
  scale_y_continuous(labels = scales::comma, n.breaks = 5) +
  scale_x_date() +
  scale_color_discrete(labels = c("Climate Change adaptation", "Climate Change mitigation", 
                                "Environmental management", "Sustainable ocean")) +
  labs(
    title = "Patents by technology sector",
    x = "Year",
    y = "NUmber of patents",
    color = "Technology domain") +
  theme_clean() +
  guides(color=guide_legend(nrow=3,byrow=TRUE)) +
  theme(legend.position = "bottom", legend.box= "vertical", legend.margin = margin())
```

```{r green-patents-domain, echo=FALSE}
#| label: green-patents-domain
#| fig-cap: Patents by Technology Sector
gp_by_tech_dom
```

Now let explore the set of independent variables that are included in the model.

## Independent variables

### Environmental Policy

I use three indicators to capture the effect of environmental policy on innovation in green technologies: The Environmental Policy Stringency Index (EPSI), the green taxes (as percentage of GDP); and the environmental R&D as a percentage of total R&D expenditure. Those three indicators capture the efforts by public authorities to create the incentives required by firms to engage in innovation in environmental technologies, and therefore, I expect them to contribute positively to national green innovative capacity.

In this section I explain in detail each of those indicators and present some basic statistics about them.

The OECD Environmental Policy Stringency Index (EPSI) used in this thesis is a country-specific and internationally-comparable measure of the stringency of environmental policy. The index is based on the degree of stringency of 13 environmental policy instruments, primarily related to climate and air pollution. The index ranges from 0 (not stringent) to 6 (highest degree of stringency) [@kruseMeasuringEnvironmentalPolicy2022]. Stringency in this context is defined as the degree to which environmental policies put an explicit or implicit price on polluting or environmentally harmful behaviour [@kruseMeasuringEnvironmentalPolicy2022].

Given the comparative nature of this thesis, the EPSI presents an important advantage. Since different countries have different policy mixes to achieve their environmental goals, some relying more on market instruments like taxes, while others focusing on non-market policies like caps and standards. By providing a synthetic tool that is more broad and comprehensive than the comparison of particular polices or policy instruments, EPSI allows for a more precise international comparison.

As of now, the EPSI includes 13 policy instruments, focusing predominantly on climate change and air pollution policies. The general index consists of “three equally-weighted subindices, which respectively group market based (e.g. taxes, permits and certificates), non-market based (e.g. performance standards) and technology support policies. Technology support policies are further divided into upstream (R&D support) and downstream (feed-in tariffs, auctions) technology support measures”. The index ranges from 0 (not stringent) to 6 (highest degree of stringency) [@MeasuringEnvironmentalPolicy2022]. 

The main limitation of the EPSI are the instruments that fall outside of its scope. For example, it does not consider policies aimed at reducing emissions from the agricultural sector [@MeasuringEnvironmentalPolicy2022]. That means that countries whose economy rely heavily on agriculture might have developed policies for that particular sector that would not be reflected in the index.

The second indicator, green taxes, is taken from the OECD Green Growth Database. It is expressed as a percentage of total tax revenue, and therefore is an indicator of the intensity of environmental taxes within the context of a country’s tax system. According to the technical documentation from the OECD, “Environmentally related taxes include taxes to (i) energy products for transport purposes (petrol and diesel) and for stationary purposes (fossil fuels and electricity); (ii) motor vehicles and transport (one-off import or sales taxes, recurrent taxes on registration or road use and other transport taxes); (iii) waste management (final disposal, packaging and other waste-related product taxes); (iv) ozone-depleting substances and (v) other environmentally related taxes.” [@oecdGreenGrowthIndicators2017].

Finally, I also include the R&D expenditure in environmental technologies. This indicator is defined as the share public R&D investment dedicated to environmental R&D activities. This indicator possibly suffers a double collinearity problem, since it is included in the general R&D considered under the general innovation infrastructure variable; and also partly considered in the EPSI indicator.

As mentioned before, the EPSI includes some green taxes and R&D expenditure as part of the instrument it considers; therefore it would not be surprising to find a strong correlation between those variables. I am including Green Taxes as a separate variable because of the potential impact that it could have not only in achieving environmental goals, but indirectly, by affecting the prices of “brown” technologies, on sparking innovation on green technologies. This straightforward mechanism is not as clear in the cases of other environmental policies. Same line of reasoning for the inclusion of environmental R&D: the direct impact in environmental innovation though a reduction of the cost of innovation by companies, makes it worth including it at least in a fist stage of the analysis. In the final model specification, I deal with the collinearity problems that might arise.

```{r, include=FALSE}

## EPSI by country in 2020

## By country

epsi_chart <- df2 %>% select(LOCATION, Year, epsi) %>% group_by(LOCATION) %>% filter(!LOCATION %in% c("LVA", "LTU")) %>% summarise(mean = mean(epsi, na.rm = TRUE)) %>% 
  ggplot() +
  geom_col(mapping = aes(x = reorder(LOCATION, mean), y = mean)) +
  scale_y_continuous(labels = scales::comma, n.breaks = 5) +
  labs(title = "EPSI, average 2000-2020",
        x = "Country",
        y = "") +
  theme_clean() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

## By year

epsi_chart2 <- df2 %>% select(Country, Year, epsi) %>% group_by(Year) %>% filter(!Country %in% c("Latvia", "Lithuania")) %>% summarise(mean = mean(epsi, na.rm = TRUE)) %>% 
  ggplot() +
  geom_line(mapping = aes(x = Year, y = mean)) +
  scale_y_continuous(labels = scales::comma, n.breaks = 5) +
  theme_clean() +
  labs(title = "EPSI, average OECD",
       x = "Year",
       y = "Average OECD",
       caption = "Note: Data for Latvia and Lithuania not available") 
```

Figure \@ref(fig:epsi-chart) Panel A shows the evolution of the environmental policy
stringency in the OECD, as measured by the Environmental Policy
Stringency Index (EPSI). In it we can see that, on average, there has
been an increase in the environmental policy stringency, which is
consistent with the increased preoccupation for effects of climate change
and environmental degradation in general. However, it is worth noting
that most of the increase happens in the fist decade of analysis, that
is from 2000 to 2010. From 2010 to 2020 the average indicator has kept
quite stable with only very minor increases.

On the other hand, Panel B shows the average score of each of the countries in the sample over the period 2000-2020. We can see that there is some heterogeneity between countries, but a majority of them are located in the 2-3 range. The highest scores are, in general, in countries in central and northern Europe (France, Denmark, Sweden), while the lower scores are in the Americas, Asia and Oceania (New Zealand, Israel, Chile, Mexico).

```{r epsi-chart, echo=FALSE}
#| label: epsi-chart
#| fig-cap: Environmental Policy Stringency Index
plot_grid(epsi_chart2, epsi_chart, nrow = 2, labels = c("A", "B"))
```

```{r env-rd, include=FALSE, echo=FALSE}
## Environmental R&D

env_rd <- df2 %>% select(Country, Year, env_rd_totalrd) %>% group_by(Year) %>% summarise(mean = mean(env_rd_totalrd, na.rm = TRUE)) %>% 
  ggplot() +
  geom_line(mapping = aes(x = Year, y = mean)) +
  scale_y_continuous(labels = scales::comma, n.breaks = 5) +
  labs(
    title = "Environmental R&D budget",
    x = "Year",
    y = "% total government R&D"
  ) +
  theme_clean()

## Environmental R&D as a % of total R&D has been decreasing in OECD countries

## By country

env_rd2 <- df2 %>% select(LOCATION, Year, env_rd_totalrd) %>% group_by(LOCATION) %>% summarise(mean = mean(env_rd_totalrd, na.rm = TRUE)) %>% 
  ggplot() +
  geom_col(mapping = aes(x = reorder(LOCATION, mean), y = mean)) +
  scale_y_continuous(labels = scales::comma, n.breaks = 5) +
  labs(title = "Environmental R&D budgets,
       average 2000-2020",
       x = "Country",
       y = "% total government R&D") +
  theme_clean() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r green-taxes, echo=FALSE, include=FALSE}

## BY Country
green_taxes_chart <- df2 %>% select(LOCATION, Year, green_taxes) %>% group_by(LOCATION) %>% 
  filter(!LOCATION %in% c("CAN", "ISR", "KOR")) %>%  summarise(mean = mean(green_taxes, na.rm = TRUE)) %>% 
  ggplot() +
  geom_col(mapping = aes(x = reorder(LOCATION, mean), y = mean)) +
  scale_y_continuous(labels = scales::comma, n.breaks = 5) +
  theme_clean() +
  labs(title = "Environmental taxes, average 2000-2020",
       x = "Country",
       y = "% of GDP",
       caption = "Note: Data not available for Canada, Israel and Korea") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

## By Year

green_taxes_chart2 <- df2 %>% select(Country, Year, green_taxes) %>% group_by(Year) %>% summarise(mean = mean(green_taxes, na.rm = TRUE)) %>% 
  ggplot() +
  geom_line(mapping = aes(x = Year, y = mean)) +
  scale_y_continuous() +
  ylim(0, 3) +
  theme_clean() +
  labs(title = "Environmental taxes, OECD average",
       x = "Country",
       y = "% of GDP")
```

Figure \@ref(fig:env-variables-1), shows the evolution of the two remaining indicators of environmental policy. Panel A shows the average share of environmental R&D budget in the general public R&D. We can see that, quite surprisingly, this share has been decreasing sharply since 2015, with only a minor increase in 2020, that only puts it at the same level it had in 2000. Panel B shows that the environmental taxes as % of GDP. This indicator has been very stable along the study period, at an average of 2.35%. However, there is small decline in the last years, going from an average of 2.44% in 2016 to 2.13% in 2020.

Just as with the EPSI, it is interesting to observe that, even as countries step up their compromises in environmental protection, that is not reflected in the indicators. It looks as if, on the contrary, environmental policy is becoming less stringent and more permissive.

```{r env-variables-1, echo = FALSE, warning=FALSE}
#| label: env-variables-1
#| fig-cap: Evolution of environmental indicators by year
plot_grid(env_rd, green_taxes_chart2, nrow = 2, labels = c("A", "B"))
```



```{r env-variables-2, echo=FALSE, warning=FALSE}
#| label: env-variables-2
#| fig-cap: Environmental indicators by country
plot_grid(env_rd2, green_taxes_chart, nrow = 2, labels = c("A", "B"))
```

### General innovation infrastructure

An important hypothesis of this thesis is that environmental policy can further environmental innovation, but also a strong innovation infrastructure is necessary, even if its effects are more mixed given the rigidities and path-dependences discussed in the theoretical framework.

To asses the general innovation infrastructure, I use a battery of six indicators, all taken from the OECD Main Science and Technology Indicators. The first is the Gross Expenditure in R&D, which measures the total expenditure on R&D from all sources, i.e. Government, business sector and higher education sector. This is the main indicator or innovation input. In a simplistic linear model of innovation [@godinLinearModelInnovation2006] it is the *only* input necessary to foster innovation. In the case of this thesis, I expect a very string positive impact on environmental innovation, but nuanced by the necessary integration of other elements in a systemic perspective.

```{r}

```


The second indicator is the share of R&D personnel per thousand total employment, which measures of the intensity of the human resources dedicated to pursuing activities related to innovation activities. Thirdly, I include the country’s investment in tertiary education as percentage of GDP. The tertiary education has a double effect on innovation: one one side, it trains the engineers and other knowledge professionals that are instrumental for the innovation processes; on the other hand, it is a major source of research, specially in basic science, which underlies many technological advancements and given its (mostly) open nature, can have large positive downstream effects. I also expect a positive impact of both of those indicators on the production of environmental technology.

Finally, I also include three indicators that are meant to capture the quality of the links between actors within the National Innovation System. The first is the share of GERD performed by the higher education sector; the second, the share of GERD financed by the business sector; and finally the share of higher education R&D financed by the business sector. Those indicators highlight the importance of the main actors in the NIS. I would expect higher links to translate to an increase in the environmental technology producing capacity. 

However, as mentioned in the theoretical framework, in general I would not be surprised to find a mixed effect of this variable on environmental innovation, with some indicators taken negative values, given that fact that a strong innovation infrastructure is needed for environmental innovation, but it might create path-dependence and lock-ins by innovation system developed on "dirty" technologies that might empower the incumbent stakeholders and hinder the development of new-to-the-world environmental technologies.


```{r innovation-infrastructure-code, echo=FALSE, include=FALSE}
rd_chart <- df2 %>%
  select(LOCATION, Year, gerd_gdp, gerd_per_hes, gerd_fin_bs, herd_by_business) %>%
  group_by(Year) %>%
  summarise(
    mean_gerd_per_hes = mean(gerd_per_hes, na.rm = TRUE),
    mean_gerd_fin_bs = mean(gerd_fin_bs, na.rm = TRUE),
    mean_herd_by_business = mean(herd_by_business, na.rm = TRUE)
  ) %>%
  ggplot(aes(x = Year)) +
  geom_line(aes(y = mean_gerd_fin_bs, color = "GERD Financed by Business Sector")) +
  geom_line(aes(y = mean_gerd_per_hes, color = "GERD Performed by Higher Education Sector")) +
  geom_line(aes(y = mean_herd_by_business, color = "HERD Performed by Business Sector")) +
  labs(
    x = "Year",
    y = "Mean Value (%)",
    color = "Indicator"
  ) +
  scale_color_manual(
    values = c("red", "blue", "steelblue"),
    labels = c("GERD Financed by Business Sector", "GERD Performed by HES", "HERD Performed by Business Sector")
  ) +
  theme_clean() +
  guides(color=guide_legend(nrow=2,byrow=TRUE)) +
  theme(legend.position = "bottom", legend.box= "vertical", legend.margin = margin())
```

```{r rd-chart, echo=FALSE}
#| label: rd-chart
#| fig-cap: Quality of linkages
rd_chart
```


### Governance Indicators

An important addition of this thesis to previous literature on environmental innovation is to include a variable on governance, that is, how the legal and political system affects the innovation system and therefore the production of environmental technologies. Measuring governance is a very challenging task in itself, since there are different ways to understand this concept by they all coincide in considering it a multidimensional and complex concept, composed of a multiplicity of elements that interact with each other. Therefore, and there is not any indicator that by itself can be used to measure it accurately. To my knowledge, the most developed effort to create index indicators that are consistent across time and comparable across countries, if the Worldwide Governance Indicators database curated by the World Bank. This database contains indicators for six dimensions: Voice and Accountability; Political Stability and Absence of Violence/Terrorism; Government Effectiveness; Regulatory Quality; Rule of Law; and Control of Corruption.

Table \@ref(tab: XXX) summarizes the definition of each of those indicators.

```{latex}
#### Table with indicators definitions
```

@kaufmannWorldwideGovernanceIndicators2010 report the aggregate measures in two ways: “in the standard normal units of the governance indicator, ranging from around ‐2.5 to 2.5, and in percentile rank terms ranging from 0 (lowest) to 100 (highest) among all countries worldwide”. For this thesis I consider only the “standard normal units” of each indicator; I also drop the margins of error, a very interesting feature of the dataset, that unfortunately cannot be properly incorporated in a regression model.

It is worth noting that OECD countries included in this study, by definition, share some characteristics of their economic and political systems. This observation means that it may be expected that most of the countries included will be relatively high in their scores, which implies that variability might be low. Chart \@ref(fig:oecd-comp-chart) confirms that observation. It shows that for 2020, on average, on the six indicators OECD countries reach positive scores, while non-OECD are on average on the negative side, with the highest differences between the two groups in control of corruption, government effectiveness, regulatory quality and rule of law. Additionally, since all six indicators are dimensions of the same conceptual construct (governance), they tend to be strongly positively correlated across countries [@kaufmannWorldwideGovernanceIndicators2010]. This strong correlation might cause problems of multicolinearity in a regression context.

```{r oecd-comparison, echo=FALSE, include=FALSE}
## Comparison of OECD and non-OECD countries

Governance_Indicators_Complete <- read.csv("G:/My Drive/EPOG/Thesis/Green_National_Innovative_Capacity/Governance_Indicators_Complete.csv")

Governance_Indicators_Complete <- Governance_Indicators_Complete %>% rename("Country" = "Country.Name", "Code" = "Country.Code", "series" = "Series.Name",
                                                                            "series_cod" = "Series.Code", "value" = "X2020..YR2020.")

Governance_Indicators_Complete <- Governance_Indicators_Complete %>% 
  mutate(OECD = if_else(Country %in% c("Australia", "Austria", "Belgium", 
                                       "Canada", "Chile", "Czech Republic", "Denmark", "Estonia", 
                                       "Finland", "France", "Germany", "Greece", "Hungary", "Iceland", 
                                       "Ireland", "Israel", "Italy", "Japan", "Korea, Rep.", "Latvia", 
                                       "Lithuania", "Luxembourg", "Mexico", "Netherlands", 
                                       "New Zealand", "Norway", "Poland", "Portugal", "Slovak Republic", 
                                       "Slovenia", "Spain", "Sweden", "Switzerland", "Türkiye", 
                                       "United Kingdom", "United States"), 1, 0))

Governance_Indicators_Complete$value <- as.numeric(Governance_Indicators_Complete$value)

gov <- Governance_Indicators_Complete %>% aggregate(value ~ series + OECD, mean, na.action = na.omit)

oecd_comp <- gov %>% group_by(series) %>% 
  ggplot() +
  geom_col(aes(x = series, y = value, fill = as.factor(OECD))) +
  labs(title = "Governance Indicators, 2020", x = "Indicator", y = "Mean Value") +
  scale_fill_brewer(breaks = c("1", "0"), labels = c("OECD", "Non-OECD")) +
  scale_x_discrete(labels = c("Control Corruption", "Government Effectiveness", "Political Stability", 
                              "Regulatory Quality", "Rule of Law", "Voice and Accountability")) +
  guides(fill=guide_legend(title="OECD")) +
  theme_clean() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r oecd-comp-chart}
#| label: oecd-comp-chart
#| fig-cap: Governance variables in OECD and non-OECD countries

oecd_comp
```

In light of that, I decided to focus on the enforcement capacity, and therefore drop the two indicators in the fist category that refers to “the process by which governments are selected, monitored and replaced”, which is not very much related to the innovation processes. Out of the two other categories, following previous literature ([@zangDoesGovernanceEfficiency2019, @dincerDoesCorruptionSlow2019, @andrijauskieneRedevelopingNationalInnovative2021]), and in order to avoid including many highly correlated indicators, I choose one of each group: quality of regulation for “the capacity of the government to effectively formulate and implement sound policies”, and control of corruption for “the respect of citizens and the State for the institutions that govern economic and social interactions among them”.

Figure \@ref(fig:gov-vars) shows the change in the values of those two variables in all OECD countries from 2000 to 2020. We can see that with very few exceptions they reach positive scores, and that those scores have not changed dramatically in the last two decades, but there has been a small average decline in the control of corruption indicator, with the average score going from 1.28 to 1.19; while in regulatory quality there has been a slight increase, from 1.18 in 2000, to 1.23 in 2020.

```{r, echo=FALSE, include=FALSE}
## Charts comparing 2000 with 2020

cont_corr_chart <- ggplot(df2, aes(x = LOCATION, y = cont_corr, fill = as.factor(Year), color = as.factor(Year))) +
  geom_col(data = subset(df2, Year == "2020"), position = "dodge", alpha = 0.5) +
  geom_point(data = subset(df2, Year == "2000"), size = 3) +
  scale_fill_manual(values = c("2020" = "steelblue"), name = "Year", labels = "2020") +
  scale_color_manual(values = c("2000" = "red"), name = "", labels = "2000") +
  labs(x = "Country", y = "Values", title = "Control of corruption") +
  theme_clean() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

reg_quality_chart <- ggplot(df2, aes(x = LOCATION, y = reg_quality, fill = as.factor(Year), color = as.factor(Year))) +
  geom_col(data = subset(df2, Year == "2020"), position = "dodge", alpha = 0.5) +
  geom_point(data = subset(df2, Year == "2000"), size = 3) +
  scale_fill_manual(values = c("2020" = "steelblue"), name = "Year", labels = "2020") +
  scale_color_manual(values = c("2000" = "red"), name = "", labels = "2000") +
  labs(x = "Country", y = "Values", title = "Regulatory Quality") +
  theme_clean() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

legend1 <- get_legend(cont_corr_chart +
    theme(legend.position = "bottom"))
```

```{r gov-vars}
#| label: gov-vars
#| fig-cap: Governance variables (2000-2020 comparison)
plot_grid(cont_corr_chart + theme(legend.position="none"), 
          reg_quality_chart + theme(legend.position="none"),
          legend1,
          nrow = 3,
          rel_heights = c(1.2, 1.2, .25)
          )
```


Finally, to asses the legal and political environment I use the World
Bank Governance Indicators, in particular quality of regulation and
perception of corruption. These indicators capture the legal, social and
political conditions that form the "environment" in which the green
innovation system has to function. I expect low corruption to be
positively associated to green national innovative capacity, since it
implies that public officials are less prone to be pressured by the
incumbent firms to maintain the status quo; and could engage in bolder
policies to support the environmental transition. I also expect better
regulation to be associated positively to green national innovative
capacity given the previous literature on the topic
[@zangDoesGovernanceEfficiency2019] and the fact that policy, including
environmental policy, requires a strong Government to enforce them. Weak
public sectors can lead to low-quality regulations that are not
well-designed and therefore become ineffective, or simply there might
not be in place adequate mechanisms for enforcement.

For more abundance, table **XX** (Appendix) shows the full list of variables with their respective
operational definition and source. Table \@ref(tab:descriptive-stats) shows the descriptive statics of all the variables.

```{r descriptive-statistics-code}
df4 <- df2 %>% select(
  Country, Year,
  green_patents, gp_mill,
  gerd_ppp, rd_tt_pt, ter_edu_gdp,  gerd_per_hes,  gerd_fin_bs,
  stock_gp, epsi, green_taxes, env_rd_totalrd,
  reg_quality, cont_corr)

descriptive_stats <- as.data.frame(df4) %>% select(
  green_patents, gp_mill,
  gerd_ppp, rd_tt_pt, ter_edu_gdp,  gerd_per_hes,  gerd_fin_bs,
  stock_gp, epsi, green_taxes, env_rd_totalrd,
  reg_quality, cont_corr)
```


```{r descriptive-stats, results='asis', echo=FALSE}
#| label: descriptive-stats
#| tab-cap: Summary statistics
sumtable(descriptive_stats, 
         labels = c("Green patents", "Green patents per million population", "GERD (PPP USD)", "R+D personnel per thousand labour force", "Public spending in tertiary education as % of GDP", "Share of GERD performed by the HES", "Share of GERD financed by the Business Sector", "Stock of green patents", "EPSI", "Green taxes as % of total tax revenue", "Environmental R+D as % of total R+D", "Regulatory Quality", "Control of corruption"), fit.page = '\\textwidth', out = "latex", title = "Descriptive Statistics")
```


## The model

The econometric model used in this study is a panel data fixed effects
model. This model accounts for both within-country and between-country
variation and controls for unobserved heterogeneity. The model is
specified as follows:

$$ 
GreenPatents_{it} = \beta_0 + \beta_1 NII_{it} + 
\beta_2 EP_{it} +
\beta_3 LPF_{it} + \alpha_i + \epsilon_{it}
$$ 

In this equation, the subscript "it" denotes the observation for country
"i" at time "t", and the variables are defined as follows:

-   GreenPatents: the number of green patents granted in country "i" at
    time "t"

-   NII: a measure of the quality of a
    country's innovation infrastructure

-   EP: a measure of a country's environmental policies

-   LPP: a measure of a country's legal and
    political factors affecting innovation and environmental protection

-   $\alpha_i$: a fixed effect capturing unobserved country-specific
heterogeneity

-   $\epsilon_{it}$: the error term for country "i" at time "t"
