```{r include_packages_2, include = FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis and also two functions
# used for labeling and referencing
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("remotes")',
        "first in the Console."
      )
    )
  }
}
if (!require(dplyr)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("dplyr", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("dplyr")',
        "first in the Console."
      )
    )
  }
}
if (!require(ggplot2)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("ggplot2", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("ggplot2")',
        "first in the Console."
      )
    )
  }
}
if (!require(bookdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("bookdown", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("bookdown")',
        "first in the Console."
      )
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
library(dplyr)
library(ggplot2)
library(knitr)
flights <- read.csv("data/flights.csv", stringsAsFactors = FALSE)
```

# Results

```{r setup, include=FALSE, echo=FALSE, warning=FALSE}

```

This section presents the main results of the analysis of the data as explained *supra*. The first part will look at where environmental innovation is happening more intensely within the OECD countries; and also presents a descriptive analysis of the main explanatory variables. Since we are both interested in the time and spatial dimension, the descriptive analysis will show the evolution of each indicator in time and also the variation across units (Countries).

Before going into it, \@ref(tab:descriptive-stats) shows the basic statistics for each of the relevant variables. This includes the two dependent variables (green patents and green patents per million population), and the battery of independent variables.

```{r, include FALSE, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
library(plotly)
library(plm)
library(pglm)
library(panelr)
library(skimr)
library(ggthemes)
library(gridExtra)
library(stargazer)
library(modelsummary)
library(kableExtra)
library(gtsummary)
library(janitor)
library(corrplot)
library(knitr)
library(tinytex)
library(huxtable)
library(vtable)

###################################################################

df2 <- read.csv("G:/My Drive/EPOG/Thesis/Green_National_Innovative_Capacity/df2.csv")

df4 <- df2 %>% select(
  Country, Year,
  green_patents, gp_mill, 
  gerd_ppp, rd_tt_pt, ter_edu_gdp,  gerd_per_hes,  gerd_fin_bs,
  stock_gp, epsi, green_taxes, env_rd_totalrd, 
  reg_quality, cont_corr)

descriptive_stats <- as.data.frame(df4) %>% select(
  green_patents, gp_mill, 
  gerd_ppp, rd_tt_pt, ter_edu_gdp,  gerd_per_hes,  gerd_fin_bs,
  stock_gp, epsi, green_taxes, env_rd_totalrd, 
  reg_quality, cont_corr)
```

```{r descriptive-stats, results='asis', echo=FALSE}
#| label: descriptive-stats
#| tab-cap: Summary statistics
sumtable(descriptive_stats, 
         labels = c("Green patents", "Green patents per million population", "GERD (PPP USD)", "R+D personnel per thousand labour force", "Public spending in tertiary education as % of GDP", "Share of GERD performed by the HES", "Share of GERD financed by the Business Sector", "Stock of green patents", "EPSI", "Green taxes as % of total tax revenue", "Environmental R+D as % of total R+D", "Regulatory Quality", "Control of corruption"), fit.page = '\\textwidth', out = "latex")
```

## Mapping Green Innovative Activity

The production of technology is general, or at least of patented
technology is very concentrated in a few countries and regions
[@oecdOECDPatentStatistics2009]. There are various reasons for that.
First, a large part of the innovative activity happen in big
laboratories located in a few countries (and even a few regions) of the
world. Second, there are differences in the patenting propensity of
firms, given the different regulations and the costs associated to
patenting in different jurisdictions.

In general, the OECD includes a sample of generally advanced countries
with some of the world's highest innovative activity and technological
endowment. Indeed, 17 out the 20 countries that filed more applications in 2020
were members of the OECD (the other three were China, India and Russia).
However, patenting activity has shifted in the last decade. China has
taken a more important role, both in terms of origin of the patents
(that is, the country of residency of the fist inventor named in the
patent filing) and the number of filings in the Chinese patent office
(CPA), both of which accounted for roughly 50% of the world's patenting
activity in 2020 [@worldintellectualpropertyorganization.WorldIntellectualProperty2021].

However, even in that sample the distribution is quite
skewed, with some countries concentrating most of the production of
technologies and others with very little patenting activity. Chart \@ref(fig:gp-by-country) shows the patenting activity in
environmental technologies of the OECD countries from 2000 to 2019. This figure shows that most of the patenting in environmental
technologies happen in six countries, the United States, Japan, Korea,
Germany, France and the United Kingdom. Out of those, the United States, Japan, ans in some years Korea, generate on average more patents than the 30 OECD countries with less patenting taken together (Others). We can also see that, in general, there was a sharp increase in the production of green patents since the beginning of the century, and especially since 2005 and until around 2012, when the tendency changes and most countries start decreasing their annual production of green patented technologies. 

```{r Fig_1_code, echo=FALSE, include=FALSE}
# 1. Green patents by country

sum_others <- df2 %>% filter(!Country %in% c("United States", "Korea", "Japan", "Germany", "France", "United Kingdom")) %>% group_by(Year) %>% filter(!Year %in% c(2020)) %>%  summarise(sum = sum(green_patents, na.rm = TRUE)) %>% mutate(Country = "Others")

gp_by_country <- df2 %>%
  filter(Country %in% c("United States", "Korea", "Japan", "Germany", "France", "United Kingdom")) %>% 
  ggplot(mapping = aes(x = Year, y = green_patents, color = Country)) +
  geom_line(linewidth = 1.2) +
  geom_line(data = sum_others, aes (x = Year, y = sum, color = Country), linewidth = 1.2) +
  labs(title = "Green patents in OECD countries 2000-2019", y = "Number of green patents", x = "Year") +
  theme_clean()

## 2. Green patents per capita

gp_per_capita <- df2 %>%
  filter(Country %in% c("United States", "Korea", "Japan", "Germany", "France", "United Kingdom")) %>% 
  ggplot(mapping = aes(x = Year, y = gp_mill, color = Country)) +
  geom_line(linewidth = 1.2) +
  labs(title = "Green patents per million population // in OECD countries 2000-2019", y = "Number of green patents", x = "Year") +
  theme_clean()

## 3. Green patents by unit of GDP

gp_per_gdp <- df2 %>%
  filter(Country %in% c("United States", "Korea", "Japan", "Germany", "France", "United Kingdom")) %>% 
  mutate(patents_gdp = green_patents / GDP) %>% 
  ggplot(mapping = aes(x = Year, y = patents_gdp, color = Country)) +
  geom_line(linewidth = 1.2) +
  labs(title = "Green patents per unit of GDP", y = "Number of green patents", x = "Year") +
  theme_clean()

```

```{r gp-by-country, warning=FALSE, echo=FALSE}
#| label: gp-by-country
#| fig-cap: Green technology patenting by country
gp_by_country
```


```{r green_patents_country, echo=FALSE, warning=FALSE}
#| label: green-patents-country
#| fig-cap: Patents by country

grid.arrange(gp_per_capita, gp_per_gdp)
```