---
title: 'Green National Innovative Capacity'
subtitle: "The drivers of Green Innovation in OECD countries"
author: 'Jeff Rodriguez'
date: 'June 2023'
institution: 'Université de Technologie de Compiègne, Sorbonne Université, Université Paris Cité, Uniiversità degli Studi di Torino'
#division: 'Mathematics and Natural Sciences'
advisor: 'Diego Useche, PhD'
# If you have more two advisors, un-silence line 7
altadvisor: 'Aldo Geuna, PhD'
altadvisor2: 'Armanda Cetrulo, PhD'
department: 'Economic Policy for the Global Transition (EPOG+)'
degree: 'Master of Science'
knit: bookdown::render_book
site: bookdown::bookdown_site

# The next two lines allow you to change the spacing in your thesis. You can 
# switch out \onehalfspacing with \singlespacing or \doublespacing, if desired.
header-includes:
    - \usepackage{setspace}\onehalfspacing
    - \usepackage{tabularray}
    - \usepackage{adjustbox}
    - \usepackage{array}
    - \usepackage{longtable}
    - \usepackage{hhline}
    - \usepackage{colortbl}
    - \usepackage{caption}
    - \usepackage{ragged2e}
    - \usepackage{amsmath}

# This will automatically install the {remotes} package and {thesisdown}
# Change this to FALSE if you'd like to install them manually on your own.
params:
  'Install needed packages for {thesisdown}': True
  
# Remove the hashtag to specify which version of output you would like.
# Can only choose one at a time.
output:
  thesisdown::thesis_pdf: default 
#  thesisdown::thesis_gitbook: default         
#  thesisdown::thesis_word: default
#  thesisdown::thesis_epub: default

# dedication: |
#   To Wes, always. I just wish you were here.

acknowledgements: |
  `r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00--prelim.Rmd")), collapse = "\n  ")`

#   This thesis would not have been possible without the support of a great lot of people:
#     
#     First and foremost, my family, without whose love and support this would have been impossible.
#     
#     My professors, both in Turin and in Paris, who inspired and gave the tools to tackle this challenge. In particular, my gratitude to Aldo Geuna for the opportunity of working and learning along him.
#     
#      Diego Useche, my main advisor, had the kindness to help me even without knowing me in advance. Thank you.
#      
#      My gratitude also to Armanda Cetrullo who gave a series of outstanding lectures and accepted to be part of my jury.
#      
#     Last but not least, my friends and classmates from EPOG. This jouney was worth it for many reasons. You are the first.

# If you are creating a PDF you'll need to write your preliminary content 
# (e.g., abstract, acknowledgements) below or use code similar to line 25-26 
# for the .RMD files. If you are NOT producing a PDF, delete or silence
# lines 25-39 in this YAML header.

abstract: |
  `r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`
#   Innovation has long been recognized as a main driver of economic growth, but more recently the challenges posed by the green transition have increased the interest in the role of innovation as a transformative force to tackle societal challenges. However, the factors that drive environmental innovation are not well understood, particularly at the national level. This thesis aims to addressing this gap in the literature by investigating the determinants of environmental innovation in OECD countries. It builds on two related streams of literature: On the one side the literature of National Systems on Innovation; and on the other, the literature on environmental innovation, both within the framework of evolutionary economics. It contributes to the literature by introducing the concept of “Green National Innovative Capacity”, understood as country’s potential to produce a steady stream of innovations in environmental-friendly technologies. Empirically, an econometric model is used to test the impact of the common innovation infrastructure; sectoral policies and the quality of governance in the production of environmental-related patents. It finds that there are significant differences in the green innovative capacity of OECD countries. The production of environmental technologies is concentrated mainly in 6 countries that account for more than half the production of green technologies in the OECD. The evidence suggests that sectoral policies (i.e. stringent environmental policy and investment in green R&D) have strong positive effects on the production of environmental technologies. The common innovation infrastructure (investment in R&D and higher education) also have a positive effect on green innovation, but their effect is nuanced by some elements of the National Innovation System that generate lock-ins and sunk costs that hinder innovation in green technologies. The evidence about the effect of the quality of governance, however, is very week and requires further research.
  
abbreviations:
  BERD: Business Sector Spending in Research and Development
  EPSI: Environmental Policy Stringency Index
  GDP: Gross Domestic Product
  GERD: Gross Domestic Spending in Research and Development
  HE: Higher Education Sector
  HERD: Higher Education Sector Spending in Research and Development
  NIC: National Innovative Capacity
  NIS: National Innovation System
  R+D: Research and Development
  SIS: Sectoral Innovation System
  SE: Standard Errors
  VIF: Variance Inflation Factor

# Note that abbreviations in lowercase letters will NOT be automatically capitalized
# Delete or silence the abbreviations section if you do not want a list of abbreviations
  
# Specify the location of the bibliography below
bibliography: Green National Innovative Capacity.bib
#link-citations: TRUE
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
always_allow_html: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line). 

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include=FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste('You need to run install.packages("remotes")",
            "first in the Console.')
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
# Set how wide the R output will go
options(width = 70)
```

<!--
The acknowledgments, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
Alternatively, you can put that content in files like 00--prelim.Rmd and
00-abstract.Rmd like done below.
-->



```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00--prelim.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00-abstract.Rmd")}

```

<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Introduction {.unnumbered}

Innovation and technological change have for a long time been recognized
as main drivers of economic growth [@solowTechnicalChangeAggregate1957;
@romerEndogenousTechnologicalChange1990;
@nelsonBringingInstitutionsEvolutionary2002;
@lundvallLearningEconomyEconomics2016]. Lately, the increasing concerns
about environmental degradation and climate change have raised the
interest in the role that innovation can play in other
areas, particularly as a transformative force to tackle societal challenges like the ongoing climate
crisis. There is a growing consensus that radical innovation is
necessary to achieve the goals set by the international comunity in terms of reduction of
emissions, and also in general for a sustainable future
[@aghionColdStartGreen2009;
@delriogonzalezEmpiricalAnalysisDeterminants2009;
@ekinsEcoinnovationEnvironmentalSustainability2010;
@gerlaghLinkingEnvironmentalInnovation2008].

However, the factors that drive environmental innovation are not well
understood, particularly at the national level. This thesis aims to
address this gap in the literature by investigating the determinants of
environmental innovation in OECD countries between 2000 and 2020.
Specifically, the thesis examines the role of common national innovation
infrastructure, sectoral (environmental) policies, and the quality of governance in
driving environmental innovation. In so doing, it builds on the
literature on National Innovation Systems
[@lundvallNationalSystemsInnovation1992], and in particular, adapting
the empirical framework of National Innovative Capacity developed by
@furmanDeterminantsNationalInnovative2002 to environmental technologies.

The main contributions of this thesis are twofold. First, in the
framework of the "twin transitions" digital and environmental, it gives
insights on the factors driving innovation in environmental technologies at the
national level. Indeed, the economic literature has explored the main
determinats of eco-innovation at the firm level
[@brunnermeierDeterminantsEnvironmentalInnovation2003;
@horbachDeterminantsEnvironmentalInnovation2008], or explored the impact
of specific environmental policies on the green innovative activity
[@francoEffectWithinSectorUpstream2017;
@gerlaghLinkingEnvironmentalInnovation2008;
@poppEnvironmentalPolicyInnovation2019]. However, there is a gap in
understanding how different policies and institutional mechanisms
interact to impact the output of environmental technologies.
Understanding those determinants is imperative to unleash the potential
of innovation to contribute to the environmental transition.

Second, this thesis employs a robust econometric model to analyze panel
data on green patents, which provides a quantitative measure of
environmental innovation. Importantly, the approach taken in this thesis
is systemic, based on the National Innovation System approach and, more
in general, on the framework of evolutionary economics
[@nelsonEvolutionaryTheoryEconomic1982;
@nelsonModernEvolutionaryEconomics2018a]. This means that it takes into
account the complex interactions between different actors, institutions,
and policies that shape innovation outcomes. By adopting this systemic
perspective, the thesis provides a more comprehensive and nuanced
understanding of the determinants of environmental innovation. This is
particularly relevant given the multi-faceted and dynamic nature of
environmental challenges, which require coordinated and adaptive policy
mixes across different levels and sectors.

For the empirical work, the thesis uses a panel data fixed effects model
on a panel OECD countries between 2000-2020. Green innovation is
measured using patent data. The main explanatory variables are the common
innovation infrastructure, sectoral (environmental) policies, and the quality of the governance. The data is sourced mainly from the
OECD, as well as other publicly
available sources.

It finds that there are significant differences in the green innovative capacity of OECD countries. The production of environmental technologies is concentrated mainly in 6 countries that account for more than half the production of green technologies in the OECD. The evidence suggests that sectoral policies (stringent environmental policy and investment in green R&D) have strong positive effects on the production of environmental technologies. The common innovation infrastructure (gross investment in R&D; investment in higher education) also have a positive effect on green innovation, but their effect is nuanced by some elements of the National Innovation System that generate lock-ins and sunk costs that hinder innovation in green technologies. The evidence about the effect of the quality of governance is however very weak and requires further research.

The thesis proceeds as follows: The next section presents the main
theoretical building blocks of the systemic approach that informs this
work. Then there is detailed description of the data and methods used
for the empirical analysis. Finally, the results are presented and
discussed.