---
title: 'Green National Innovative Capacity'
author: 'Jeff Rodriguez'
date: 'June 2023'
institution: 'Université de Technologie de Compiègne, Sorbonne Université, Université Paris Cité, Uniiversità degli Studi di Torino'
#division: 'Mathematics and Natural Sciences'
advisor: 'Diego Useche, PhD'
# If you have more two advisors, un-silence line 7
altadvisor: 'Aldo Geuna, PhD'
department: 'Economics'
degree: 'Master of Science'
knit: bookdown::render_book
site: bookdown::bookdown_site

# The next two lines allow you to change the spacing in your thesis. You can 
# switch out \onehalfspacing with \singlespacing or \doublespacing, if desired.
header-includes:
    - \usepackage{setspace}\onehalfspacing

# This will automatically install the {remotes} package and {thesisdown}
# Change this to FALSE if you'd like to install them manually on your own.
params:
  'Install needed packages for {thesisdown}': True
  
# Remove the hashtag to specify which version of output you would like.
# Can only choose one at a time.
output:
  thesisdown::thesis_pdf: default 
#  thesisdown::thesis_gitbook: default         
#  thesisdown::thesis_word: default
#  thesisdown::thesis_epub: default

# If you are creating a PDF you'll need to write your preliminary content 
# (e.g., abstract, acknowledgements) below or use code similar to line 25-26 
# for the .RMD files. If you are NOT producing a PDF, delete or silence
# lines 25-39 in this YAML header.
abstract: |
  `r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab 
# is needed on the line after the `|`.
acknowledgements: |
  I want to thank a few people.
dedication: |
  You can have a dedication here if you wish. 
preface: |
  This is an example of a thesis setup to use the reed thesis document class 
  (for LaTeX) and the R bookdown package, in general.
abbreviations:
  # ABC: American Broadcasting Company
  # CBS: Colombia Broadcasting System
  # CUS: Computer User Services
  # PBS: Public Broadcasting System
  # NBC: National Broadcasting Company
# Note that abbreviations in lowercase letters will NOT be automatically capitalized
# Delete or silence the abbreviations section if you do not want a list of abbreviations
  
# Specify the location of the bibliography below
bibliography: Green National Innovative Capacity.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
always_allow_html: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line). 

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include=FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste('You need to run install.packages("remotes")",
            "first in the Console.')
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
# Set how wide the R output will go
options(width = 70)
```

<!--
The acknowledgments, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
Alternatively, you can put that content in files like 00--prelim.Rmd and
00-abstract.Rmd like done below.
-->



```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00--prelim.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00-abstract.Rmd")}

```

<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Introduction {.unnumbered}

Innovation and technological change have for a long time been recognized
as main drivers of economic growth [@solowTechnicalChangeAggregate1957;
@romerEndogenousTechnologicalChange1990;
@nelsonBringingInstitutionsEvolutionary2002;
@lundvallLearningEconomyEconomics2016]. However the increasing concerns
about environmental degradation and climate change have raised the
interest in the role innovation can play in other areas, particularly as
a transformative force to tackle societal challenges like the climate
crisis. There is a growing consensus that radical innovation is
necessary to achieve the goals set by the world in terms of reduction of
emissions, and also in general for a sustainable future
[@aghionColdStartGreen2009;
@delriogonzalezEmpiricalAnalysisDeterminants2009;
@ekinsEcoinnovationEnvironmentalSustainability2010;
@gerlaghLinkingEnvironmentalInnovation2008].

However, the factors that drive environmental innovation are not well
understood, particularly at the national level. This thesis aims to
address this gap in the literature by investigating the determinants of
environmental innovation in OECD countries between 2000 and 2020.
Specifically, the thesis examines the role of national innovation
infrastructure, environmental policy, and legal and political factors in
driving environmental innovation. In so doing, it builds on the
literature on National Innovation Systems
[@lundvallNationalSystemsInnovation1992], and in particular, adapting
the empirical framework of National Innovative Capacity developed by
@furmanDeterminantsNationalInnovative2002 to environmental technologies.

The main contributions of this thesis are twofold. First, in the
framework of the "twin transitions" digital and environmental, it gives
insights on the factors driving green innovative capacity at the
national level. Indeed, the economic literature has explored the main
determinats of eco-innovation at the firm level
[@brunnermeierDeterminantsEnvironmentalInnovation2003;
@horbachDeterminantsEnvironmentalInnovation2008], or explored the impact
of specific environmental policies on the green innovative activity
[@francoEffectWithinSectorUpstream2017;
@gerlaghLinkingEnvironmentalInnovation2008;
@poppEnvironmentalPolicyInnovation2019]. However, there is a gap in
understanding how different policies and institutional mechanisms
interact to impact the output of environmental technologies.
Understanding those determinants is imperative to unleash the potential
of innovation to contribute to the environmental transition.

Second, this thesis employs a robust econometric model to analyze panel
data on green patents, which provides a quantitative measure of
environmental innovation. Importantly, the approach taken in this thesis
is systemic, based on the National Innovation System approach and, more
in general, on the framework of evolutionary economics
[@nelsonEvolutionaryTheoryEconomic1982;
@nelsonModernEvolutionaryEconomics2018a]. This means that it takes into
account the complex interactions between different actors, institutions,
and policies that shape innovation outcomes. By adopting this systemic
perspective, the thesis provides a more comprehensive and nuanced
understanding of the determinants of environmental innovation. This is
particularly relevant given the multi-faceted and dynamic nature of
environmental challenges, which require coordinated and adaptive policy
mixes across different levels and sectors.

For the empirical work, the thesis uses a panel data fixed effects model
on a panel OECD countries between 2000-2020. Green innovation is
measured using patent data. The main explanatory variables are national
innovation infrastructure, environmental policy, and legal and political
factors. Negative Binomial and Poisson regression models are also used
to test the robustness of the findings. The data is sourced from the
OECD Green Growth Indicators Databse, as well as other publicly
available sources.

[FINDINGS]

The thesis proceeds as follows: The next section presents the main
theoretical building blocks of the systemic approach that informs this
work. Then there is detailed description of the data and methods used
for the empirical analysis. Finally, the results are presented and
discussed.